Smart Drill-Down: Interactive Summarization of Relational Tables

### 2014-11-21 TableSummarization/Paper/TableSummarization.tex
small changes
Addressed Comments. Added NP-Hardness proof
added new graphs to writeup

%\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\pdfpagewidth=8.5in
\pdfpageheight=11in

@@

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

@@

\title{Interactive Summarization of Relational Tables}

\author{
Manas Joglekar\\Stanford University\\manasrj@stanford.edu
\and
Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
\and
Aditya Parameswaran\\Stanford University\\adityagp@cs.stanford.edu
}
\begin{document}
\maketitle

\begin{abstract}
In this paper we present {\em smart drill down},
a system for interactively exploring a relational table
to discover ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a $1000$ tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively drill down on an existing rule to
explore that part of the table.
We describe an
algorithm for finding the approximately optimal list of rules, and an optimized
sampling scheme for efficiently summarizing large tables. Finally, we
demonstrate the usefulness of smart drill down by performing experiments on
a real dataset.
\end{abstract}

>>>>

\usepackage{amsmath}
\makeatletter
\let\@copyrightspace\relax
\makeatother
\pdfpagewidth=8.5in
\pdfpageheight=11in

\usepackage{enumitem}
\usepackage{times}
\usepackage{subfigure}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,color}
\usepackage{verbatim}
\usepackage{framed}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{framed}
\usepackage[normalem]{ulem}

@@

\renewcommand*\ttdefault{cmvtt}
\usepackage[T1]{fontenc}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{example}[definition]{Example}
\newcounter{prob}
\newtheorem{problem}[prob]{Problem}

\newcommand{\agp}[1]{\textcolor{red}{Aditya: #1}}
\newcommand{\mrj}[1]{\textcolor{green}{#1}}
\newcommand{\agpdel}[1]{\textcolor{blue}{\bf ||Del: #1||}}

@@

\title{Smart Drill-Down: Interactive Summarization of Relational Tables}

\author{
Manas Joglekar\\Stanford University\\manasrj@stanford.edu
\and
Hector Garcia-Molina\\Stanford University\\hector@cs.stanford.edu
\and
Aditya Parameswaran\\University of Illinois\\adityagp@illinois.edu
}
\begin{document}
\maketitle

\begin{abstract}
In this paper we present {\em smart drill down},
a system for interactively exploring a relational table
to discover ``interesting'' groups of tuples.
Each group of tuples is described by a {\em rule}.
For instance, the rule $(a, b, \star, 1000)$ tells us that
there are a $1000$ tuples with value $a$ in the first column and $b$
in the second column (and any value in the third column).
Smart drill down presents an analyst with a list of rules that
together describe interesting aspects of the table.
The analyst can tailor the definition of interesting,
and can interactively drill down on an existing rule to
explore that part of the table.
We describe an
algorithm for finding the approximately optimal list of rules, and an optimized
sampling scheme for efficiently summarizing large tables. Finally, we
demonstrate the usefulness of smart drill down by performing experiments on
real datasets.
\end{abstract}

### 2014-11-21 TableSummarization/Paper/exp.tex
added new graphs to writeup

\subsection{Dataset}
The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

@@

Unless otherwise specified, in all our experiments, we restrict the table to the first $7$ columns in order to make the result tables fit into the page. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.

\subsection{Qualitative Study}
We first perform a qualitative study of our Smart Drill Down system. We observe the effects of various user interface operations on our dataset, and then try out a few different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

\subsection{Quantitative Study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm.

@@

The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule. Thus for size scoring, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$. For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$. At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively. Of course, the minimum value of $m_w$ we can use is not known to us beforehand. But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.

@@

The computation time increases approximately linearly with parameter $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.

In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}. This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

>>>>

\subsection{Datasets}
The first dataset, denoted `Marketing', contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

@@

The second dataset, denoted `Census', is a US 1990 Census dataset from the UCI Machine Learning repository~\cite{uciml}. It consists of about $2.5$ million tuples, with each tuple corresponding to a person. It has $68$ columns, including ancestry, age, and citizenship. Numerical columns, such as age, have been bucketized beforehand in the dataset. We use this dataset in the experiments in Section~\ref{sec:quantitative_study} in order to study the accuracy and performance of sampling on a large dataset.

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit into the page. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting function, and $20$ for the Bits weighting function.

\subsection{Qualitative Study}
We first perform a qualitative study of our Smart Drill Down system. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

% TODO: Add the experiment where you compare to vanilla a priori.

\subsection{Quantitative Study}\label{sec:quantitative_study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm. We use both the Marketing and Census datasets. The Marketing dataset is relatively small with around $9000$ tuples, whereas the Census dataset is quite large, with $2.5$ million tuples. We observe that the accuracy and running time of our algorithm depends largely on the $m_w$ and $minSS$ parameters, rather than on the underlying database size.

@@

The value of $m_w$ required to ensure a correct answer is equal to the maximum weight of a selected rule. Thus for size scoring on the Marketing dataset, according to Table~\ref{table:uiexample1}, we require $m_w \geq 3$. For the second weighting function, according to Table~\ref{table:weigtingbitwise}, the minimum required value of $m_w$ is $10$. At these values of $m_w$, we see that the expansion takes $1.5$ seconds and about $0.25$ seconds respectively. Of course, the minimum value of $m_w$ we can use is not known to us beforehand. But even if we use more conservative values of $m_w$, say $6$ and $20$ respectively, the running times are about $1.5$ and $0.5$ seconds respectively.

@@

The computation time increases approximately linearly with parameter $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.
% TODO redo the census bits weighting experiments with proper m_w (forgot to set it correctly earlier)

% TODO, find for census dataset.
In addition, we measure the number of incorrect rules per iteration. If the correct set of rules to display is $r_1, r_2, r_3$ and the displayed set is $r_1, r_3, r_4$ then that means there is one incorrect rule. We find the number of incorrect displayed rules across $50$ iterations, and display the average value in Figure~\ref{fig:minSS_error_rule}. This number is almost always $0$ for the Size weighting function, and between $1$ and $2$ for the Bits weighting function. Note that even when we display an `incorrect' rule, it is usually the $5^{th}$ or $6^{th}$ best rule instead of one of the top $4$ rules, which still results in a reasonably good summary of the table.

### 2014-11-21 TableSummarization/Paper/intro.tex
New graphs. Yet to be added to writeup
added new graphs to writeup

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by another aggregate,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the system displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

@@

\item Based on our implemented smart drill down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on a real dataset,
and show additional examples of smart drill down in action.

>>>>

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by a Sum aggregate over some measure column,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the system displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

@@

\item Based on our implemented smart drill down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on real datasets,
and show additional examples of smart drill down in action.
